---
title: "Lab 1"
author: "Brady Lamson"
format: pdf
editor: visual
---

# Section A

```{r}
murder <- c(15.1, 11.3, 7.8, 10.1, 10.3, 6.8, 3.1, 6.2, 10.7, 13.9)
illiteracy <- c(2.1, 1.5, 1.8, 1.9, 1.1, 0.7, 1.1, 0.9, 1.3, 2.0)
```

```{r, include=FALSE}
my.reg <- lm(murder~illiteracy)
my.reg |> summary()
```


## Part 5

The estimate of $\beta_0$ is $b_0 = 1.982$

The estimate of $\beta_1$ is $b_1 = 5.242$

The estimated standard error $b_1$ i $s\{b_1\} = 1.828$

The observed value of the test statistic for the t test of

$$
H_0: \beta_1 = 0 \\
H_a: \beta_1 \neq 0
$$

is $t = 2.867$

The $\text{p-value}= .02$

Is the observed $b_1$ statistically significantly different from zero (Yes/No)? 

Yes

What does this indicate about whether there's any relationship between the murder rate and the illiteracy rate?

There appears to be a positive relationship between murder and illiteracy rate.

$R^2 = .51$

Based on the value of $R^2$, what percentage of the variation in murder rates is explained by illiteracy rate? 

$51\%$

## Part 6

Does the assumption of normally distributed errors appear to be met? 

Yes.

## Part 7

Does the assumption of a constant standard deviation assumption appear to be met?

Yes.

## Part 8

### a

$$
\begin{aligned}
SSR &= 60 \\
SSE &= 58.413 \\
MSR &= 60.008 \\
MSE &= 7.302 \\
SSTO &= 118.421 \\
F &= 8.2183
\end{aligned}
$$

### b

Give the degrees of freedom for SSR: 1

Give the degrees of freedom for SSSE: 8

Give the degrees of freedom for SSTO: 9

### c

Show that F is equal to the t test statistic from step 5.

```{r}
2.867^2
```

Both p-values are the same, .0209 and .0209.

```{r}
sse <- 58.413
ssto <- 118.421
1 - (sse / ssto)
```

.5067 is approximately .51, which, as I rounded, are essentially the same.

# Section B

```{r, include=FALSE}
my.reg <- lm(murder ~ 1)
summary(my.reg)
```

## Part 2

The estimate of $\beta_0$ is $b_0 = 9.530$

The estimated standard error $b_1$ i $s\{b_0\} = 1.147$

As no $R^2$ value shows up here it is 0. As such the $R^2$ is larger for the model that includes illiteracy as a predictor.

## Part 3

The line is horizontal. This makes sense as the predictor of 1 assumes that the rate of murder is constant, unchanging.

## Part 4

```{r}
mean(murder)
```

9.53 is the same as the estimate for $b_0$.

## Part 5

```{r, include=FALSE}
t.test(x=murder, mu=0)
```

Are the results the same as the test for $\mu$? Yes.

# Section C

```{r, include=FALSE}
my.reg <- lm(murder ~ illiteracy)
confint(my.reg)
```

## Part 2
The interval for $\beta_1$ is $(1.02, 9.46)$. This means that the slope could be anywhere between 1.02 and 9.46.

## Part 3

```{r, include=FALSE}
new_dat <- data.frame(illiteracy = 1.6)
predict(my.reg, newdata = new_dat)
```

The value of the estimate of the mean response is 10.367

## Part 4
```{r}
predict(my.reg, newdata = new_dat, interval = "confidence")
```

This says that the mean murder of the new data could be anywhere between 8.26 and 12.45.

## Part 5

```{r}
predict(my.reg, newdata = new_dat, interval = "prediction")
```

This predicts that the murder rate of the new data is between 3.79 and 12.94