---
title: "Homework 1"
author: "Brady Lamson"
format: pdf
editor: visual
---

# Problem 1.5

When asked to state the simple linear regression model, a student wrote it as follows:

$$
E\{Y_i\} = \beta_0 + \beta_1X_i + \epsilon_i
$$

Do you agree?

**Answer**

No. The thing that ruins their answer is that it is giving the expected value of $Y_i$. That changes things. The book on page nine points out that:

$$
E\{Y_i\} = \beta_0 + \beta_1X_i
$$

Note the lack of an epsilon at the end. 

# 1.7 

In a simulation exercise, regression model applies with $\beta_0=100$, $\beta_1=20$ and $\sigma^2 = 25$. An observation on Y will be made for $X=5$.

## Part A

Can you state the exact probability that Y will fall between 195 and 205? Explain? 

No. Not unless I know the distribution of the error is normal.

## Part B

Yes. 

$$
\begin{aligned}
E(\hat{Y}) &= \beta_0 + \beta_1 X_i = 100 + 20(5) = 200 \\
\sigma &= \sqrt{25} = 5
\end{aligned}
$$

```{r}
bounds <- c(195,205)
mean <- 200
sd <- 5

# This creates a vector of two probabilities, one from (-inf,195) 
# and another from (-inf, 205). 
probs <- pnorm(bounds, mean, sd)

# Subtract upper bound prob from lower bound prob
# Equivalent to normalcdf(195,205,)
prob <- probs[2] - probs[1]
```

```{r, echo=FALSE}
glue::glue("The probability the predicted value falls between ({bounds[1]}, {bounds[2]}) is approximately {round(prob, 3)}")
```


# 1.18 

$\sum e_i = 0$ when regression model 1.1 is fitted to a set of n cases by the method of least squares. Is it also true that $\sum \epsilon_i = 0$? 

I would assume that no, it doesn't apply. As $\epsilon$ is an unknown batch of random variables that are independent and uncorrelated so the sum of these values is also random and unknown. 

# 1.19

```{r}
data_pr19 <- read.table("CH01PR19.txt", header = T)
reg_19 <- lm(gpa ~ act.score, data=data_pr19)
reg_19 |> summary()
```

## Part A

$$
\begin{aligned}
\beta_0 &= 2.114 \\
\beta_1 &= 0.038 \\ 
\hat{Y} &= 2.114 + 0.038X \\
\end{aligned}
$$

## Part B

```{r}
plot(data_pr19$act.score, data_pr19$gpa, main = "ACT Scores vs. GPA", pch=20)
abline(reg = reg_19, col="red", lwd = 3)
```

This line doesn't appear to fit the data *well* and the $R^2$ is only $0.07$ which indicates some large deviation away from the line. Regardless, there does seem to be a noticable though slight positive relationship here.

## Part C

```{r}
predict(reg_19, newdata = data.frame(act.score=30))
```

## Part D
```{r}
predict(reg_19, newdata = data.frame(act.score=31))
```

# 1.21

```{r}
data_pr21 = read.table("CH01PR21.txt", header = T)
reg_21 <- lm(ampules ~ transfers, data=data_pr21)
reg_21 |> summary()
```

## Part A and Part D

$$
\hat{Y} = 10.2 + 4X
$$

```{r}
plot(x=data_pr21$transfers, y=data_pr21$ampules, pch=20, main="Transfers vs. Ampules")
abline(reg_21, col="blue", lwd=3)
points(x=mean(data_pr21$transfers), y=mean(data_pr21$ampules), col = "red", pch=20, cex=3)
```

The line appears to be a good fit with a high $R^2$ of 0.9. 

## Part B

```{r}
new_data = data.frame(transfers = 1)
predict(reg_21, new_data)
```

## Part C

Going from 1 transfer to 2 transfers it is expected for the number of broken ampules to increase by 4.

## Part D

The red dot on the plot above from part B represents $(\overline{X}, \overline{Y})$. As the line crosses through the line this has been verified.

# 1.23 

## Part A

```{r}
reg_19$residuals |> sum()
```

This only doesn't show 0 due to a computational rounding error. This is essentially 0.

## Part B

```{r}
sd_reg <- reg_19$residuals |> sd()
var_reg <- sd_reg^2
c(sd_reg, var_reg)
```

$\sigma$ is expressed in units of the response variable, which in this case is destroyed ampules. 

# 1.25

## Part A

```{r}
reg_21$residuals[1]
```

1.8 is the first residual but after reading what I could I still really do not understand the relationship between this and the first random error $\epsilon_1$. 

## Part B

```{r}
reg_21$residuals^2 |> sum()
```

```{r}
reg_21 |> anova()
```

$$
\begin{aligned} 
\sum e_i^2 = 17.6 \\
MSE = 2.2
\end{aligned}
$$

MSE estimates the variance of the residuals.