---
title: "Lab 4"
author: "Brady Lamson"
format: pdf
editor: visual
---

```{r, include=FALSE}
data <- read.table("../../datasets/satisfaction.txt", header = T)
```

# Part A

```{r}
pairs(data)
```

## 3

```{r, include=FALSE, echo=FALSE}
cor(data)
```

Satisfaction is most highly correlated with age with a correlation of -.787

Highest negative correlation is satisfaction and age, highest positive correlation is severity and anxiety with a correlation of .670.

## 4

```{r, include=FALSE, echo=FALSE}
model <- lm(satisfaction ~ age + severity + anxiety, data= data)

model |> summary()
```

### A

$$
\hat{Y} = 158.49 -1.14X_1 -0.44X_2 - 13.47X_3
$$

The value of $b_1$ indicates that satisfaction drops as age increases.

The value of $b_2$ indicates that satisfaction slightly drops as severity increases. Though this result is not statistically significant.

The value of $b_3$ indicates that satisfaction drastically drops as severity increases. Though this result is not statistically significant.


### B

$$R^2 = 0.68$$

68% of the variation in satisfaction is explained by the model with age, severity and anxiety.

### C

$$
F = 30.05\\
p = 1.542 \cdot 10^{-10}
$$

As the p-value is far lower than any reasonable alpha, we have evidence to reject the null hypothesis that all of the slopes are equal to 0. There is a significant relationship between at least one of the predictors.

### D

$$\sigma^2 = 10.06^2 = 101.2036\\
\sigma=10.06$$

### E

$$
s\{b_1\} = .2148\\
s\{b_2\} = .4920\\
s\{b_3\} = 7.10\\
$$

### F

B1 test:

$$
t = -5.315 \\
p = 3.81 \cdot 10^{-6} \\
yes
$$

B2 test:

$$
t = -.898 \\
p = .3741 \\
no
$$

B3 test:

$$
t = -1.897 \\
p = .0647 \\
no
$$

## 5

I would say the assumption of normality of the error term appears to be reasonable. 

## 6.

Yup, constant standard deviation seems to be okay here.

# Part B

```{r, include=FALSE, echo=FALSE}
lengthCm <- c(60, 69, 66, 64, 54, 67, 59, 65, 63)
weightG <- c(136, 198, 194, 140, 93, 172, 116, 174, 145)
lengthIn <- c(24.0, 27.6, 26.4, 25.6, 21.6, 26.8, 23.6, 26.0, 25.2)
diameterCm <- c(2.1, 2.2, 2.2, 2.2, 1.6, 2.2, 2.0, 2.1, 2.2)
snaaaaaaaaaaaaaaakes <- data.frame(LengthCm = lengthCm,
LengthIn = lengthIn,
DiameterCm = diameterCm,
WeightG = weightG)
```

```{r, include=FALSE, echo=FALSE}
my.reg <- lm(WeightG ~ LengthCm + LengthIn + DiameterCm, data = snaaaaaaaaaaaaaaakes)

my.reg |> summary(0)
```

LengthCm and LengthIn are totally dependent on each other, they're the exact same measurement. 

# Part C

```{r, echo=FALSE, include=FALSE}
Site <- c(1, 2, 3, 4)
CO2_Flux <- c(1.242, 0.590, 1.677, 1.090)
SoilTemp <- c(11.35, 13.55, 7.85, 11.50)
AirTemp <- c(23.0, 16.8, 24.3, 16.2)
PctSand <- c(69.6, 73.6, 60.2, 73.1)
PctSilt <- c(19.0, 16.3, 29.9, 19.7)
PctClay <- c(11.4, 10.1, 9.9, 7.2)
rm_soil <- data.frame(CO2_Flux, SoilTemp, AirTemp, PctSand, PctSilt, PctClay)
```

There isnt enough rows of data to compute these values. Too many predictors, not enough data. 